---
title: "Detentions, post- de-duplication"
author: "Cindy J. Pang"
date: "2026-03-04"
format:
  pdf:
    documentclass: article
    fontsize: 11pt
    geometry: margin=1in
    number-sections: false
    colorlinks: true
    keep-tex: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
rm(list=ls()) 
```

```{r}
library(dplyr)
library(tidyr)
library(arrow)
library(ggplot2)
library(lubridate)
```

# Compare Dropped Counts to Original
```{r}
# only read-in a few rows of the detentions data to avoid storing the whole thing and create flags for which rows from the original to drop  
detentions_temp <- read_feather(here::here("data/ice-processed/detentions-merged-with-flags.feather"), col_select = c("Detention_Book_In_Date", "source_file", "drop_row"))|>
  mutate(
    week_start = floor_date(Detention_Book_In_Date, unit = "week", week_start = 7),
    row_id = row_number()
  )|>
    arrange(week_start)|>
    filter(drop_row == FALSE)

detentions_counts_original <- read_feather(here::here("data/ice-counts/detentions-weekly-counts.feather"))|>
  rename(n0 = n)

weekly_counts <- detentions_temp |>
  group_by(week_start) |>
  summarise(
    n = n(),
    n_source_files = n_distinct(source_file),
    source_files = paste(sort(unique(source_file)), collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(week_start)

compare_counts <- merge(weekly_counts, 
  detentions_counts_original, 
  by = c("week_start"), 
  all = TRUE)|>
  mutate(
    n_dropped = n0-n
  )

# Visual Inspection of Counts - counts look like they mostly overlap
ggplot(compare_counts, aes(x=week_start))+
  geom_line(aes(y=n0, color = "Original Count"), alpha = 0.5)+
  geom_line(aes(y=n, color = "Dropped Count"), alpha = 0.5)+
  labs(
    x = "Week", 
    y = "Counts", 
    title = "Comparison of Dropped Counts vs. Original Counts"
  )+
  theme_minimal()
```

The dropped counts and original seem to mostly overlap (no big concerns/issues). The time period between 2015-2020 seem to have aggresive drops in counts.  
```{r}
ggplot(compare_counts, aes(x=week_start, y=n_dropped))+
  geom_line()+
  labs(
    x = "Week", 
    y = "Counts", 
    title = "# dropped counts"
  )+
  theme_minimal()
```

# Plot of Detentions (post-duplication drops)  
```{r}
df <- compare_counts |>
  arrange(week_start) |>
  mutate(
    source_files_norm = ifelse(is.na(source_files), "", source_files),
    changed = source_files_norm != lag(source_files_norm, default = first(source_files_norm)),
    seg_id = cumsum(changed),
    seg_code = paste0("S", seg_id)
  )

# segment metadata (for label placement)
segments <- df |>
  group_by(seg_id, seg_code) |>
  summarise(
    seg_start = min(week_start),
    seg_end   = max(week_start),
    seg_mid   = seg_start + as.numeric(seg_end - seg_start)/2,
    source_files = first(source_files_norm),
    .groups = "drop"
  )

# vlines at segment starts (exclude first segment)
vlines <- segments |>
  arrange(seg_start) |>
  slice(-1)

# y position for segment codes (near top)
y_top <- max(df$n0, df$n_dropped, na.rm = TRUE)

ggplot(df, aes(x = week_start)) +
  geom_line(aes(y = n), linewidth = 0.5) +
  geom_vline(
    data = vlines,
    aes(xintercept = as.numeric(seg_start)),
    linetype = "dashed",
    alpha = 0.6,
    inherit.aes = FALSE
  ) +
  geom_label(
    data = segments,
    aes(x = seg_mid, y = y_top, label = seg_code),
    vjust = -0.2,
    size = 3,
    label.size = 0.2,
    alpha = 0.95,
    inherit.aes = FALSE
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +
  labs(
    title = "Weekly Removals with Data Source Change Points",
    subtitle = "See Segment Codes (SX) to see time periods",
    x = "Week start",
    y = "Counts",
    color = NULL
  ) +
  theme_minimal()

segments_print <- segments |>
  select(seg_code, seg_start, seg_end, source_files)
knitr::kable(segments_print, col.names = c("Segment Code", "Start Date", "End Date", "Source Files"))
```

# Weekly Detentions by Source File  
```{r}
weekly_counts_by_src <- detentions_temp |>
  group_by(source_file, week_start) |>
  summarise(
    n = n(),
    .groups = "drop"
  ) |>
  arrange(source_file, week_start)

ggplot(weekly_counts_by_src, 
       aes(x = week_start, y = n, color = source_file)) +
  geom_line(size = 0.5, alpha = 0.5) +
  labs(
    title = "Weekly Detention Counts by Source File",
    x = "Week Start",
    y = "Number of Detentions",
    color = "Source File"
  ) +
  theme_minimal()
```

It looks like the only files we need would be `2023_ICFO_42034` and `120125` past a certain point.  

## Drop other source files for final dataset  
```{r}
detentions_data_final <- detentions_temp |>
  mutate(keep_row = case_when(
    source_file == "2024-ICFO-41855" & Detention_Book_In_Date < as.Date("2009-09-27") ~ TRUE, 
    source_file == "uwchr" & Detention_Book_In_Date >= as.Date("2009-09-27") & Detention_Book_In_Date < as.Date("2023-12-31")~ TRUE, 
    source_file == "120125" & Detention_Book_In_Date >= as.Date("2023-12-31") ~ TRUE, 
    TRUE ~ FALSE
  ))

weekly_counts_by_src <- detentions_data_final |>
  filter(keep_row == TRUE)|>
  group_by(source_file, week_start) |>
  summarise(
    n = n(),
    .groups = "drop"
  ) |>
  arrange(source_file, week_start)

ggplot(weekly_counts_by_src, 
       aes(x = week_start, y = n, color = source_file)) +
  geom_line(size = 0.5, alpha = 0.5) +
  labs(
    title = "Weekly Detainer Counts by Source File",
    x = "Week Start",
    y = "Number of Detentions",
    color = "Source File"
  ) +
  theme_minimal()
```

```{r, eval=F}
rows_to_keep <- detentions_data_final |>
  filter(keep_row == TRUE)|>
  select(row_id)

write_feather(rows_to_keep, "code/detentions-cpang/detentions_rows_to_drop.feather")
```



