---
title: "Missing Counts Summary for ICE Data"
author: "Cindy J. Pang"
date: "2026-02-04"
format:
  pdf:
    documentclass: article
    fontsize: 11pt
    geometry: margin=1in
    number-sections: false
    colorlinks: true
    keep-tex: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
rm(list=ls())
```

```{r}
library(arrow)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(here)
```

```{r}
all_data <- read_feather(here("data", "ice-counts", "all-events-by-week.feather")) |>
  select(week_start, event_type, count) |>
  mutate(Year = year(week_start))
```

## Arrests - `Apprehension_Date` Counts  

### By-Week
```{r}
all_data |>
  subset(event_type == "arrests") |>
  ggplot(aes(x = week_start, y = count)) +
  geom_col(fill = "black", color = "black", linewidth = 0.2) +
  scale_x_date(expand = c(0, 0)) +
  theme_minimal()+
  labs(
    title = "Arrests by Week", 
    x = "Week", 
    y = "Counts"
  )
```

Subset to 2010 onwards to see what is going on. 

```{r}
all_data |>
  filter(event_type == "arrests", week_start >= as.Date("2010-01-01")) |>
  ggplot(aes(x = week_start, y = count)) +
  geom_col(fill = "grey70", color = "black", linewidth = 0.2) +
  scale_x_date(expand = c(0, 0)) +
  theme_minimal()+
  labs(
    title = "Arrests from 2010 Onwards", 
    x = "Week", 
    y = "Counts"
  )
```

### By-Year 
```{r}
all_data |>
  subset(event_type == "arrests") |>
  ggplot(aes(x = Year, y = count)) +
  geom_col(fill = "black", color = "black", linewidth = 0.2) +
  theme_minimal()+
  labs(
    title = "Arrests by Year", 
    x = "Year", 
    y = "Counts"
  )
```


## Detentions - `Detention_Book_In_Date` Counts  

### By-Week
```{r}
all_data |>
  subset(event_type == "detentions") |>
  ggplot(aes(x = week_start, y = count)) +
  geom_col(fill = "grey70", color = "black", linewidth = 0.2) +
  scale_x_date(expand = c(0, 0)) +
  theme_minimal()+
  labs(
    title = "Detentions by Week", 
    x = "Week", 
    y = "Counts"
  )
```

### By-Year 

```{r}
all_data |>
  subset(event_type == "detentions") |>
  ggplot(aes(x = Year, y = count)) +
  geom_col(fill = "black", color = "black", linewidth = 0.2) +
  theme_minimal()+
  labs(
    title = "Detentions by Year", 
    x = "Year", 
    y = "Counts"
  )
```

## Detainers  

### By-Week 
```{r}
all_data |>
  subset(event_type == "detainers") |>
  ggplot(aes(x = week_start, y = count)) +
  geom_col(fill = "grey70", color = "black", linewidth = 0.2) +
  scale_x_date(expand = c(0, 0)) +
  theme_minimal()+
  labs(
    title = "Detainers by Week", 
    x = "Week", 
    y = "Counts"
  )
```

### By-Year 

```{r}
all_data |>
  subset(event_type == "detainers") |>
  ggplot(aes(x = Year, y = count)) +
  geom_col(fill = "black", color = "black", linewidth = 0.2) +
  theme_minimal()+
  labs(
    title = "Detainers by Year", 
    x = "Year", 
    y = "Counts"
  )
```

## Encounters  

### By-Week 
```{r}
all_data |>
  subset(event_type == "encounters") |>
  ggplot(aes(x = week_start, y = count)) +
  geom_col(fill = "grey70", color = "black", linewidth = 0.2) +
  scale_x_date(expand = c(0, 0)) +
  theme_minimal()+
  labs(
    title = "Encounters by Week", 
    x = "Week", 
    y = "Counts"
  )
```

### By-Year 

```{r}
all_data |>
  subset(event_type == "encounters") |>
  ggplot(aes(x = Year, y = count)) +
  geom_col(fill = "black", color = "black", linewidth = 0.2) +
  theme_minimal()+
  labs(
    title = "Encounters by Year", 
    x = "Year", 
    y = "Counts"
  )
```

By inspection we are **missing data** from `2023-02-05`-`2023-08-20`. 

## Removals  

### By-Week 
```{r}
all_data |>
  subset(event_type == "removals") |>
  ggplot(aes(x = week_start, y = count)) +
  geom_col(fill = "grey70", color = "black", linewidth = 0.2) +
  scale_x_date(expand = c(0, 0)) +
  theme_minimal()+
  labs(
    title = "Removals by Week", 
    x = "Week", 
    y = "Counts"
  )
```

### By-Year 

```{r}
all_data |>
  subset(event_type == "removals") |>
  ggplot(aes(x = Year, y = count)) +
  geom_col(fill = "black", color = "black", linewidth = 0.2) +
  theme_minimal()+
  labs(
    title = "Removals by Year", 
    x = "Year", 
    y = "Counts"
  )
```

## Summary of Data Missingness  

| Event Type | Years Covered | Missing Gaps |
|------------|---------------|--------------|
| Arrests    | 2011-Oct. 12, 2025     | Jan. 2011-Sept. 18|
| Detentions | 1983-Oct. 12, 2025     |              |
| Detainers  | 1998-Mar 16, 2025     | 1999-2000, 2022-2004, 2006, 2008, 2012; Mar. 23,2025-Oct. 12, 2025 |
| Encounters | 2011-2025     | Jan. 2011 - Sept. 18, 2011; Feb. 2023 - Aug 2023; Aug. 2025-Oct. 2025 |
| Removals   | 2013-Jul. 27, 2025 | Feb. 5, 2023-Aug. 20, 2023; Aug. 3, 2025-Oct. 12, 2025 |

## Appendix 

### By Year  

```{r}
# Arrests 
arrests_yearly <- all_data |>
  filter(event_type == "arrests", count > 0) |>
  mutate(Year = year(week_start)) |>
  group_by(Year) |>
  summarise(
    total_arrests = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(Year)

# Detainers
detainers_yearly <- all_data |>
  filter(event_type == "detainers", count > 0) |>
  mutate(Year = year(week_start)) |>
  group_by(Year) |>
  summarise(
    total_detainers = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(Year)

# Detentions 
detentions_yearly <- all_data |>
  filter(event_type == "detentions", count > 0) |>
  mutate(Year = year(week_start)) |>
  group_by(Year) |>
  summarise(
    total_detention_BookIns = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(Year)

# Encounters 
encounters_yearly <- all_data |>
  filter(event_type == "encounters", count > 0) |>
  mutate(Year = year(week_start)) |>
  group_by(Year) |>
  summarise(
    total_encounters = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(Year)

# Removals 
removals_yearly <- all_data |>
  filter(event_type == "removals", count > 0) |>
  mutate(Year = year(week_start)) |>
  group_by(Year) |>
  summarise(
    total_removals = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(Year)
```
```{r}
years_all <- Reduce(function(x, y) full_join(x, y, by = "Year"),
                     list(arrests_yearly, detainers_yearly, detentions_yearly, encounters_yearly, removals_yearly))|>
                      arrange(Year)
```
```{r}
knitr::kable(years_all, col.names = c("Year", "Arrests", "Detainers", "Detentions", "Encounters", "Removals"))
```

### Inspect Weekly Patterns for Data-Missingness  

```{r}
arrests_weekly_past2011 <- all_data |>
  filter(event_type == "arrests", week_start >= as.Date("2011-01-01")) |>
  group_by(week_start) |>
  summarise(
    total_arrests = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(week_start)

detainers_weekly_past1998 <- all_data |>
  filter(event_type == "detainers", week_start >= as.Date("1998-01-01")) |>
  group_by(week_start) |>
  summarise(
    total_arrests = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(week_start)

detentions_weekly_past1983 <- all_data |>
  filter(event_type == "detentions", week_start >= as.Date("1983-01-01")) |>
  group_by(week_start) |>
  summarise(
    total_detentions = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(week_start)

encounters_weekly_past2011<- all_data |>
  filter(event_type == "encounters", week_start >= as.Date("2011-01-01")) |>
  group_by(week_start) |>
  summarise(
    total_encounters = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(week_start)

removals_weekly_past2013 <- all_data |>
  filter(event_type == "encounters", week_start >= as.Date("2013-01-01")) |>
  group_by(week_start) |>
  summarise(
    total_removals = sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(week_start)

knitr::kable(arrests_weekly_past2011, col.names = c("Week", "Arrests"), caption = "Arrests Past 2011")
knitr::kable(detainers_weekly_past1998, col.names = c("Week", "Detainers"), caption = "Detainers Past 1998")
knitr:: kable(detentions_weekly_past1983, col.names = c("Week", "Detention Book Ins"), caption = "Detentions Past 1983")
knitr::kable(encounters_weekly_past2011, col.names = c("Week", "Encounters"), caption = "Encounters Past 2011")
knitr::kable(removals_weekly_past2013, col.names = c("Week", "Removals"), caption = "Removals Past 2013")
```




